### 2020-05-11 Revision_2.Rmd
version 1

affiliation:
  - id            : "1"
    institution   : "Utrecht University faculty of Social and Behavioral Sciences, department of Methodology & Statistics"
  - id            : "2"
    institution   : "Erasmus University Rotterdam Erasmus School of Social and Behavioral Sciences, department of Public Administration and Sociology"

abstract: |
  This longitudinal study of Australian families (n = 1884, from age 6-12)
  examined how fathers' and mothers' quantitative involvement (time spent on
  childcare) and qualitative involvement (specific parenting behaviors) relate
  to children's emotional adjustment development. We used dynamic structural
  equation modeling to disentangle stable between-family differences from
  within-family fluctuations in qualitative parenting and emotional adjustment,
  allowing us to investigate the direction of effects between parents and
  children. Because fathers have been theorized to contribute uniquely to
  emotional adjustment development, we examined differences between mothers and
  fathers, and the interplay between parent- and child sex. We further examined
  whether between-family differences in quantitative involvement,
  operationalized as joint (total) and relative (one parent does more)
  involvement, predict qualitative parenting behaviors and emotional
  adjustment, and moderate the within-family interplay between them. Our
  results primarily revealed between-family correlations of qualitative
  involvement with emotional adjustment. Evidence for within-family effects was
  limited to a parenting effect of maternal warmth, and several child effects,
  which indicated that emotional adjustment evokes changes in parenting.
  Parenting effects varied substantially across families however, especially
  for fathers. Between-family differences in quantitative involvement explained
  some of these differences. When joint quantitative involvement was higher,
  both parents engaged in more desirable qualitative parenting. When mothers
  were relatively more involved than fathers, fathers displayed less desirable
  qualitative parenting, and children were less well-adjusted. There was some
  limited evidence that within-family predictive effects were moderated by
  quantitative involvement, particularly relative involvement. These findings
  suggest that all family members benefit from a more equal childcare
  distribution.

@@

figsintext        : no
figurelist        : no
tablelist         : no
footnotelist      : no
lineno            : yes
mask              : yes

@@

\vspace{\baselineskip}
Children's emotional adjustment is defined as the absence of internalizing symptoms and emotional dysregulation [@goodmanPsychometricPropertiesStrengths2001]. It is integral to social competence, academic achievement, and mental health [@grossHandbookEmotionRegulation2013]. Ample research has demonstrated associations between children's emotional adjustment and parenting; both in terms of quantitative involvement (the amount of time parents spend caring for their child), and in terms of qualitative parenting behaviors (how parents behave towards their child). Nevertheless, the existing literature has two major shortcomings. Firstly, although it is widely acknowledged that quantitative and qualitative involvement are both relevant to child development [@faganShouldResearchersConceptualize2014; @pleckPaternalInvolvementRevised2004], these dimensions are rarely jointly investigated, let alone in interaction [@craigDoesFatherCare2006; @faganShouldResearchersConceptualize2014]. Yet it is plausible that quantitative involvement might moderate predictive effects between qualitative parenting and emotional adjustment. For example, children's emotional adjustment might be differently impacted by anger coming from a highly involved parent, than from a generally absent parent. Conversely, children's poor emotional adjustment might elicit warm responses from hands-on parents, who are aware of the reasons for their children's emotional troubles, than from out-of-touch parents, who might perceive maladjustment as a problem behavior that should be disciplined with anger or inductive reasoning. The present study overcomes this limitation by investigating both qualitative and quantitative involvement, and the interplay between them, in relation to children's emotional adjustment.

Secondly, although emotional adjustment is one particular area of development where fathers have been theorized to play a unique role [@paquetteTheorizingFatherchildRelationship2004], fathers are still not consistently included in research [@faganShouldResearchersConceptualize2014]. Consequently, prior research has not sufficiently addressed mothers' and fathers' joint and unique contributions to emotional adjustment development. The present study overcomes this limitation by incorporating both parents. We use sophisticated methods to examine how joint and relative quantitative involvement at age 6 moderate the longitudinal interplay between mothers' and fathers' qualitative parenting and children's emotional adjustment, from the age of 6 to 12 years. We thus obtain a more comprehensive and nuanced understanding of how fathers' and mothers' involvement relate to children's emotional adjustment development.

@@

Early developmental research focused on parental socialization; the notion that parents shape child development through proximal mechanisms, such as qualitative parenting behaviors [@maccobySocializationContextFamily1983; see also @vanlissaRoleFathersMothers2019]. Qualitative parenting behaviors are related to the dimensions underlying parenting styles: Responsiveness and demandingness [@maccobySocializationContextFamily1983]. Substantial evidence indicates that responsiveness, which we operationalized in terms of warm and supportive parenting, is positively associated with children's emotional adjustment [@morrisRoleFamilyContext2007]. Therefore we hypothesized that parental warmth would be positively associated with children's emotional adjustment, $H_{1a}$.<!--@khalequePerceivedParentalAcceptanceRejection2002; -->

The literature linking demandingness and emotional adjustment is less conclusive [@morrisRoleFamilyContext2007]. Demandingness encompasses parenting behaviors related to setting and enforcing expectations for children. Some of these behaviors help children meet societal expectations, whereas others are overly restrictive, and interfere with children's autonomy needs [@barberCentralityControlParenting2013]. We examined two aspects of demandingness: Consistency; the extent to which consequences for transgressions are clear and enforced, and inductive reasoning; explaining the reasons for rules and punishments. Consistency correlates positively with emotional adjustment [@brandPerceivedParentalInconsistency1990]. Therefore we also hypothesized that consistency would be positively associated with emotional adjustment, $H_{1b}$. There is less empirical evidence for links between inductive reasoning and emotional adjustment. Given that inductive reasoning promotes the internalization of parental expectations, including display rules for emotions [@grusecImpactParentalDiscipline1994], we hypothesized that it would also be positively related to emotional adjustment, $H_{1c}$.

Negative parenting, which we operationalized as parental anger, is a qualitative parenting dimension not originally subsumed in the parenting styles typology. It is negatively linked with children's emotional adjustment [@changHarshParentingRelation2003; @eisenbergParentalReactionsChildren1999]. Parental anger negatively impacts children's emotional adjustment [@scaramellaParentingChildrenDistress2008], because it is upsetting, and because anger reflects poor parental emotion regulation, which serves as a model for children [@changHarshParentingRelation2003]. We thus hypothesized that anger would be negatively related to emotional adjustment, $H_{1d}$.

@@

The family is commonly viewed as a dynamic system, whose members reciprocally influence one another [@kerrPerceivedParentingStyle2012]. Nevertheless, bi-directional effects are not routinely investigated [@flouriFathersBehaviorsChildren2010]. Most of the research cited in support of hypotheses $H_{1a-d}$ is cross-sectional, and could also support child effects. We thus hypothesize that emotional adjustment positively predicts warmth, consistency, and inductive reasoning, and negatively predicts anger, $H_{2a-d}$. These predictions are in line with studies that investigated child effects specifically, and found that maladjustment elicits less warmth and more negative parenting [@gadeyneLongitudinalRelationsParenting2004]. There is also evidence for bi-directional positive links between parental consistency and emotional adjustment [@lenguaBidirectionalAssociationsTemperament2005]. The adolescent literature provides additional evidence that emotional adjustment positively predicts parental support [@otterpohlCrossLaggedRelationsParenting2015; @vandergiessenPerceivedAutonomySupport2014], and negatively predicts negative parenting [@wijsbroekDirectionEffectsPerceived2011]. Thus, in a way, parents respond positively to emotional adjustment, and negatively to maladjustment.
<!-- Neppl 2009: Adolescents' harsh behavior towards parents predicts harsh parenting, and positive behavior predicts positive parenting -->

@@

Methods commonly used to examine direction of effects between parents and children have been criticized for failing to disentangle between-family differences from within-family predictive effects [@hamakerCritiqueCrosslaggedPanel2015]. The recently developed random-intercept cross-lagged panel model, which is an extension of the standard cross-lagged panel model [@hamakerCritiqueCrosslaggedPanel2015], and dynamic structural equation modeling (DSEM); a multilevel generalization of this model [@asparouhovDynamicStructuralEquation2018], overcome this limitation by disentangling the two levels.

Several studies have re-examined the direction of effects between parents and children using these methods, but mostly in adolescent samples. One such study found that effects of parenting on adjustment that emerged using classical methods became non-significant when stable between-family differences were taken into account [@keijsersParentalMonitoringAdolescent2016]. Another study similarly found that associations between maternal and paternal qualitative parenting and adolescents' emotional adjustment, were mostly explained by between-family level correlations. The remaining predictive effects consisted primarily of child effects, not parenting effects [@vanlissaRoleFathersMothers2019]. The authors theorized that this finding was reflective of the relative equality of parent-child relationships in adolescence. If this is correct, we would expect to find more evidence of parenting effects in the present, younger, sample. These two studies also illustrate that applying these new methods changes the pattern of results in unexpected ways. Consequently, it is impossible to make specific predictions about the pattern of between-family correlations and within-family predictive effects. For now, we will assume that $H_{1a-d}$ and $H_{2a-d}$ apply at both the between- and within-family level.

@@

The notion that fathers should share child-rearing responsibilities with mothers has become widely accepted in the Western world [@pleckFatherhoodIdealsUnited1997]. Consequently, researchers have come to study qualitative involvement along the same dimensions for mothers and fathers [@maccobySocializationContextFamily1983; @pleckPaternalInvolvementRevised2004]. Fathers' qualitative involvement is associated with emotional adjustment, even after accounting for mothers [@flouriFathersBehaviorsChildren2010]. Some research suggests that both parents influence children similarly, but that effects might be stronger for mothers [@lambDevelopmentSignificanceFatherChild2004; @faganShouldResearchersConceptualize2014]. Others have argued that mothers and fathers might contribute through somewhat different mechanisms. Attachment relationships are grounded in a balance between warmth and security, and exploration [@grossmannDeveloppementAttachementAdaptation1998]. Some have proposed that warmth is more central to mother-child attachment relationships, whereas father-child relationships can be construed as "activation relationships", with greater emphasis on exploration [@paquetteFatherChildActivationRelationship2013]. By encouraging children to venture outside their comfort zone, fathers stimulate a broad spectrum of emotions, thereby expanding children's emotional repertoire, and engaging their developing regulatory abilities  [@flouriFathersBehaviorsChildren2010; @bogelsDoesFatherKnow2011; @parkeFatherInvolvement2000].

In line with this theory, empirical research has found that mothers' responses to emotional expression are more warm and supportive than fathers' responses, which are more demanding and punitive [@cassanoInfluenceGenderParental2007; @eisenbergParentalReactionsChildren1999]. In adolescence, maternal warmth was found to be a more salient socialization mechanism of emotional adjustment than maternal demandingness, whereas this pattern was reversed for fathers [@vanlissaRoleFathersMothers2019]. Similar differentiated associations were found for younger children's psychological adjustment [@mcdowellDifferencesMothersFathers2003]. Finally, there is initial evidence that emotional adjustment is more negatively impacted by mothers' negative parenting, compared to fathers [@changHarshParentingRelation2003]; perhaps because negative parenting is somewhat incongruous with the warm "attachment" relationship typically expected of mothers. Based on this research, we therefore hypothesize that emotional adjustment would be more strongly associated with maternal warmth and anger, and with paternal demandingness, operationalized as consistency and inductive reasoning, $H_{3}$.

@@

Early fatherhood research primarily operationalized involvement in *quantitative* terms: Father absence in divorced families, and childcare sharing in dual-earner families [@lambBiosocialPerspectivePaternal1987]. Quantitative involvement has been researched extensively in terms of the absolute amount of childcare provided [see @pleckPaternalInvolvementRevised2004]. However, some have argued that the *division* of childcare tasks is more relevant for children's adjustment [e.g., @keizerInfluenceFathersMothers2019; @craigHowMothersFathers2011]. Indeed, the equity of the childcare distribution is associated with socio-emotional development [@deutschPaternalParticipationChild2001]. Moreover, an unequal distribution often implies that women perform a disproportionate share of the work [@craigDoesFatherCare2006]. This strains the marital relationship [@shockleyItNotWhat2018], and erodes parenting quality, particularly for parents of difficult children [@sheltonInterparentalConflictNegative2008].

@@

By the same reasoning, we might hypothesize that greater differential involvement would positively predict the desirable parenting behaviors of the more involved parent, and negatively predict those of the less involved parent, $H_{5a-d}$ [@sabattiniRelationMothersFathers2004]. However, there are also reasons to expect greater differential involvement to be negatively associated with both parents' parenting quality, and with children's emotional adjustment. Retrospective reports indicate that an unequal childcare distribution is associated with suboptimal parenting behaviors: Increased authoritarian parenting in mothers, and increased disengaged parenting in fathers [@sabattiniRelationMothersFathers2004]. One reason might be that parents are less able to monitor each other's child-rearing behavior when they are not equally involved [@mclanahanConsequencesFatherAbsence2014]. Additionally, differential involvement might negatively impact the marital relationship, and thereby undermine both parents' involvement quality. Childcare responsibilities are a heavy burden [@deutschPaternalParticipationChild2001]. An unfair distribution strains the interparental relationship, interferes with sensitive parenting, and undermines child well-being [@davies2006interparental]. When care is distributed relatively equally, mothers tend to be more satisfied [@pharesFathersMothersInvolvement2009], and the family system more harmonious. Consequently, balanced co-parenting is associated with greater parental warmth [@barnettInterdependenceParentingMothers2008].

@@

Bioecological theories argue that children require a certain amount of care; who provides it is somewhat flexible [see @bronfenbrennerEcologyHumanDevelopment1979]. We therefore hypothesize that greater joint care should positively predict children's emotional adjustment, $H_{6a}$. Further, it has been argued that children feel less supported when they perceive that one of their parents is relatively uninvolved, than when both parents are equally invested in their well-being [@deutschPaternalParticipationChild2001]. We therefore predict that greater relative involvement will negatively predict emotional adjustment, $H_{6b}$.

@@

Interactions between involvement quality and quantity have largely been overlooked in the parenting literature [@jaffeeLifeFatherBenefits2003]. Consequently, it is not possible to substantiate specific hypotheses for such interactions. One notable exception is Feinberg's coparenting model [@feinbergInternalStructureEcological2003]. Coparenting refers to the ways parents relate to each other in the parenting role. Although the coparenting literature is typically invoked to describe child-rearing in divorced families, it also holds insights for intact families. According to Feinberg, the "division of labor" in childcare, analogous to our operationalization of relative involvement, affects qualitative child-rearing behaviors and children's adjustment directly, and moderates the link between these two constructs. Applying this insight from the coparenting literature, we might expect parents' quantitative involvement to moderate effects between qualitative parenting behaviors and emotional adjustment. Due to the limited literature, it is not possible to formulate specific hypotheses about the nature of these moderating effects. There is, however, preliminary evidence to indicate that interactions between involvement quality and quantity are relevant. For instance, fathers' playful and caring involvement positively predicted preschoolers' emotional adjustment in school, but only when accompanied by a supportive coparenting relationship [@jiaTransactionalRelationsFather2012]. Furthermore, the quantity of paternal school-related involvement interacted with paternal warmth to positively predict adolescents' academic achievement and behavioral adjustment [@wangParentalInvolvementAfrican2014]. We thus explored the moderating effects of joint and relative quantitative involvement on the interplay between qualitative parenting behaviors and emotional adjustment.

@@

The present study examined whether specific qualitative parenting behaviors are associated with children's emotional adjustment, and whether these associations are moderated by quantitative involvement. We aimed to make several contributions. First, by including both mothers and fathers, we can tease apart their common and unique roles in children's emotional adjustment development. Second, by incorporating both qualitative and quantitative measures of involvement, we can examine interactions between qualitative and quantitative involvement, which have been largely overlooked in prior work. Third, by disentangling stable between-family differences from dynamic within-family predictive processes, we can establish at which level associations occur, and assess the within-family direction of effects between qualitative involvement and emotional adjustment. Finally, we examined child sex as a moderator of between-family differences in all study variables and within-family predictive effects.

@@

We employed data from the `r masked("K-cohort")` of the `r masked("Longitudinal Study of Australian Children (LSAC)")`, waves `r masked("2-5")`, which were collected by means of biannual home visits, and self-report mail questionnaires, from 2006-2012. Each wave was reviewed and approved by the
`r masked("Australian Institute of Family Studies Ethics Committee")`, which is a Human Research
Ethics Committee registered with the `r masked("National Health and Medical Research Council")`. Documentation of the ethics approval is available online [@australianinstituteoffamilystudiesGrowingAustralia2020]. `r masked("LSAC")` employed a clustered sampling design to obtain a representative sample of the `r masked("Australian")` population. Complete details regarding the study design, sample, and procedure are available online [@australianinstituteoffamilystudiesGrowingAustralia2020].

@@

We calculated mean scores for all study variables with multiple indicators. Scale descriptive statistics, number of items, attrition (which ranged from 7-14% across respondents), reliability, and intraclass correlations (ICCs) are reported in Supplementary Table 1; scale correlations are provided in Supplementary Table 2. The difference in fit between structurally invariant models and metrically invariant models was smaller than .01 on all conventional fit indices: RMSEA, CFI, and TLI (see Supplementary Table 3). We thus deemed it defensible to assume metric measurement invariance (i.e., to assume that the factor loadings of each construct are the same across time points and respondents).

@@

Responsiveness and inductive reasoning were measured using the warmth (e.g., "How often do you express affection by hugging, kissing and holding this child?") and inductive reasoning (e.g., "Talk it over and reason with this child when he/she misbehaved?") subscales of the Child Rearing Questionnaire [CRQ, @patersonAssociationBehaviouralAdjustment2001]. Consistency and negative parenting were measured using the consistency (e.g., "How often is this child able to get out of punishment when he/she really sets his/her mind to it?", reverse coded) and anger (e.g., "How often are you angry when you punish this child?") scales developed by the Canadian National Longitudinal Survey of Children and Youth [NLSCY; @statisticscanadaNationalLongitudinalSurvey2000]. These self-report instruments used five-point scales, ranging from "never or almost never" to "always or almost always".

@@

Measures of quantitative involvement varied from wave to wave, due to children's changing care needs [@wangATUSFathersInvolvement2009]. We used the first wave reports as an indicator of between-family differences in quantitative involvement. Both parents rated their involvement in four different child care activities on a five-point scale ranging from "once a day or more" to "not at all" [e.g., "In the past month how often did you help this child brush his/her teeth?"; @baxterFatheringAustraliaCouple2011].

@@

Emotional adjustment was measured using the reverse-coded emotional distress subscale of the teacher-reported Strengths and Difficulties Questionnaire, on a three-point scale from "not true" to "certainly true" [e.g., "Often unhappy, downhearted or tearful"; @goodmanPsychometricPropertiesStrengths2001].

>>>>

affiliation:
  - id            : "1"
    institution   : "Utrecht University, department of Methodology & Statistics"
  - id            : "2"
    institution   : "Erasmus University Rotterdam Erasmus School of Social and Behavioral Sciences, department of Public Administration and Sociology"

abstract: |
  This longitudinal study of Australian families (n = 1884, from age 6-12)
  examined how fathers' and mothers' quantitative involvement (time spent on
  childcare) and qualitative involvement (specific parenting behaviors) relate
  to children's emotional adjustment development. We used dynamic structural
  equation modeling to disentangle stable between-family differences from
  within-family fluctuations in qualitative parenting and emotional adjustment,
  allowing us to investigate the direction of effects between parents and
  children. Because fathers have been theorized to contribute uniquely to
  emotional adjustment development, we examined differences between mothers and
  fathers, and the interplay between parent- and child sex. We further examined
  whether between-family differences in quantitative involvement,
  operationalized as joint (total) and relative (one parent does more)
  involvement, predict qualitative parenting behaviors and emotional
  adjustment, and moderate the within-family interplay between them. Results
  revealed between-family correlations of qualitative
  involvement with emotional adjustment. Evidence for significant average
  within-family effects was
  limited to a parenting effect of maternal warmth, and several child effects,
  which indicated that emotional adjustment evokes changes in parenting.
  Parenting effects varied substantially across families however, especially
  for fathers. Between-family differences in quantitative involvement explained
  some of these differences. When joint quantitative involvement was higher,
  both parents engaged in more desirable qualitative parenting. When mothers
  were relatively more involved than fathers, fathers displayed less desirable
  qualitative parenting, and children were less well-adjusted. There was some
  limited evidence that within-family predictive effects were moderated by
  quantitative involvement, particularly relative involvement. These findings
  suggest that all family members benefit from a more equal childcare
  distribution.

@@

figsintext        : no
figurelist        : no
tablelist         : no
footnotelist      : no
lineno            : yes
mask              : no
urlcolor          : blue

@@

\vspace{\baselineskip}
Emotional adjustment is defined as the absence of internalizing symptoms and emotional dysregulation [@goodmanPsychometricPropertiesStrengths2001]. It is integral to social competence, academic achievement, and mental health [@grossHandbookEmotionRegulation2013]. Ample research has demonstrated associations between children's emotional adjustment and parental involvement; both quantitative involvement (the amount of time parents spend caring for their child), and qualitative involvement (specific parenting behaviors). Nevertheless, the existing literature has two major shortcomings. Firstly, although it is widely acknowledged that quantitative and qualitative involvement are both relevant [@faganShouldResearchersConceptualize2014; @pleckPaternalInvolvementRevised2004], these dimensions are rarely investigated together, let alone in interaction [@craigDoesFatherCare2006; @faganShouldResearchersConceptualize2014]. The present study overcomes this limitation by investigating both qualitative and quantitative involvement, and the interplay between them, in relation to children's emotional adjustment.

Secondly, although fathers have been theorized to play a unique role in emotional adjustment development [@paquetteTheorizingFatherchildRelationship2004], they are not consistently included in research [@faganShouldResearchersConceptualize2014]. Consequently, little is known about mothers' and fathers' joint and unique contributions to emotional adjustment development. The present study overcomes this limitation by incorporating both parents. We use sophisticated methods to examine how joint and relative quantitative involvement at age 6 moderate the longitudinal interplay between mothers' and fathers' qualitative parenting and children's emotional adjustment, from the age of 6 to 12 years. We thus obtain a more comprehensive and nuanced understanding of how fathers' and mothers' involvement relate to children's emotional adjustment development.

@@

Early developmental research focused on parental socialization; the notion that parents shape child development through qualitative parenting dimensions, such as responsiveness and demandingness [@maccobySocializationContextFamily1983]. Substantial evidence indicates that responsiveness, which we operationalized in terms of warm and supportive parenting behaviors, is positively associated with children's emotional adjustment [@morrisRoleFamilyContext2007]. Therefore we hypothesized that parental warmth would be positively associated with children's emotional adjustment, $H_{1a}$.<!--@khalequePerceivedParentalAcceptanceRejection2002; -->

The literature linking demandingness and emotional adjustment is less conclusive [@morrisRoleFamilyContext2007]. Demandingness encompasses parenting behaviors related to rule-setting and enforcement. Some of these behaviors help children meet societal expectations, whereas others are overly restrictive, and interfere with children's autonomy needs [@barberCentralityControlParenting2013]. We examined two aspects of demandingness: Consistency; the extent to which consequences for transgressions are clear and enforced, and inductive reasoning; explaining the reasons for rules and punishments. As prior research found positive correlations between consistency and emotional adjustment [@brandPerceivedParentalInconsistency1990], we also hypothesized to find these positive associations, $H_{1b}$. There is less evidence for links between inductive reasoning and emotional adjustment. Given that inductive reasoning promotes the internalization of parental expectations, including display rules for emotions [@grusecImpactParentalDiscipline1994], we hypothesized that it would also be positively related to emotional adjustment, $H_{1c}$.

Finally, negative parenting, which we operationalized as parental anger, is a qualitative parenting behavior negatively assosicated with children's emotional adjustment [@changHarshParentingRelation2003; @eisenbergParentalReactionsChildren1999]. Parental anger negatively impacts children's emotional adjustment [@scaramellaParentingChildrenDistress2008], because it is upsetting, and because anger reflects poor parental emotion regulation, which serves as a model for children [@changHarshParentingRelation2003]. We thus hypothesized that anger would be negatively related to emotional adjustment, $H_{1d}$.

@@

Contemporary theory views the family as a dynamic system, whose members reciprocally influence one another [@kerrPerceivedParentingStyle2012]. Nevertheless, child effects on parents are not routinely investigated [@flouriFathersBehaviorsChildren2010]. As most of the research cited in support of hypotheses $H_{1a-d}$ is cross-sectional, it could also support child effects. We thus hypothesize that emotional adjustment positively predicts warmth, consistency, and inductive reasoning, and negatively predicts anger, $H_{2a-d}$. These predictions are in line with studies that specifically investigated child effects, which found that maladjustment elicits less warmth and more negative parenting [@gadeyneLongitudinalRelationsParenting2004]. There is also evidence for bi-directional positive links between parental consistency and emotional adjustment [@lenguaBidirectionalAssociationsTemperament2005]. The adolescent literature provides additional evidence that emotional adjustment positively predicts parental support [@otterpohlCrossLaggedRelationsParenting2015; @vandergiessenPerceivedAutonomySupport2014], and negatively predicts negative parenting [@wijsbroekDirectionEffectsPerceived2011]. This suggests that parents respond positively to emotional adjustment, and negatively to maladjustment.
<!-- Neppl 2009: Adolescents' harsh behavior towards parents predicts harsh parenting, and positive behavior predicts positive parenting -->

@@

Methods commonly used to examine direction of effects between parents and children have been criticized for failing to disentangle between-family differences from within-family predictive effects [@hamakerCritiqueCrosslaggedPanel2015]. This limitation is addressed by the random-intercept cross-lagged panel model [@hamakerCritiqueCrosslaggedPanel2015], and the dynamic structural equation modeling (DSEM); a multilevel generalization of this model [@asparouhovDynamicStructuralEquation2018]. Several studies have re-examined the direction of effects between parents and children using these methods, but mostly in adolescent samples. One such study found that effects of parenting on adjustment that emerged using classical methods became non-significant when stable between-family differences were taken into account [@keijsersParentalMonitoringAdolescent2016]. Another study similarly found that associations between maternal and paternal qualitative parenting and adolescents' emotional adjustment, were mostly explained by between-family level correlations. The remaining predictive effects consisted primarily of child effects, not parenting effects [@vanlissaRoleFathersMothers2019]. The authors theorized that this finding was reflective of the relative equality of parent-child relationships in adolescence. If this is correct, we should expect more evidence of parenting effects in the present, younger, sample. However, these prior studies demonstrate that applying these new methods changes the pattern of results in unexpected ways. Consequently, it is impossible to make specific predictions about the pattern of between-family correlations and within-family predictive effects. For now, we will assume that $H_{1a-d}$ and $H_{2a-d}$ apply at both the between- and within-family level.

@@

The notion that fathers should share child-rearing responsibilities with mothers has become commonplace in the Western world [@pleckFatherhoodIdealsUnited1997]. Consequently, researchers have come to study qualitative involvement along the same dimensions for mothers and fathers [@maccobySocializationContextFamily1983; @pleckPaternalInvolvementRevised2004]. Fathers' qualitative involvement is associated with emotional adjustment, even after accounting for mothers [@flouriFathersBehaviorsChildren2010]. Some research suggests that both parents influence children similarly, but that effects might be stronger for mothers [@lambDevelopmentSignificanceFatherChild2004; @faganShouldResearchersConceptualize2014]. Others have argued that mothers and fathers might contribute through somewhat different mechanisms. Attachment relationships are grounded in a balance between warmth and security, and exploration [@grossmannDeveloppementAttachementAdaptation1998]. Some have proposed that warmth is more central to mother-child attachment relationships, whereas father-child relationships can be construed as "activation relationships", with greater emphasis on exploration [@paquetteFatherChildActivationRelationship2013]. By encouraging children to venture outside their comfort zone, fathers stimulate a broad spectrum of emotions, which expands children's emotional repertoire and exercises their regulatory abilities  [@flouriFathersBehaviorsChildren2010; @bogelsDoesFatherKnow2011; @parkeFatherInvolvement2000].

In line with this theory, research has found that mothers' respond more warmly and supportively to emotional expression, whereas fathers are more demanding and punitive [@cassanoInfluenceGenderParental2007; @eisenbergParentalReactionsChildren1999]. In adolescence, maternal warmth was a more salient socialization mechanism of emotional adjustment than maternal demandingness, whereas this pattern was reversed for fathers [@vanlissaRoleFathersMothers2019]. Similar differentiated associations were found for younger children's psychological adjustment [@mcdowellDifferencesMothersFathers2003]. Finally, there is initial evidence that emotional adjustment is more negatively impacted by mothers' negative parenting, compared to fathers [@changHarshParentingRelation2003]; perhaps because negative parenting is somewhat incongruous with the warm "attachment" relationship typically expected of mothers. Based on this research, we hypothesize that emotional adjustment would be more strongly associated with maternal warmth and anger, and with paternal demandingness, operationalized as consistency and inductive reasoning, $H_{3}$.

@@

Early fatherhood research primarily operationalized involvement in *quantitative* terms: Father absence in divorced families, and childcare sharing in dual-earner families [@lambBiosocialPerspectivePaternal1987]. Quantitative involvement has been researched extensively in terms of the absolute amount of childcare provided [see @pleckPaternalInvolvementRevised2004]. However, some have argued that the *division* of childcare tasks is more relevant for children's adjustment [e.g., @keizerInfluenceFathersMothers2019; @craigHowMothersFathers2011]. The equity of the childcare distribution is associated with socio-emotional development [@deutschPaternalParticipationChild2001]. Moreover, an unequal distribution often implies that women perform a disproportionate share of the work [@craigDoesFatherCare2006]. This strains the marital relationship [@shockleyItNotWhat2018], and erodes parenting quality, particularly for parents of difficult children [@sheltonInterparentalConflictNegative2008].

@@

By the same reasoning, we hypothesize that greater differential involvement will positively predict the desirable parenting behaviors of the more involved parent, and negatively predict those of the less involved parent, $H_{5a-d}$ [@sabattiniRelationMothersFathers2004]. However, there are also reasons to expect that greater differential involvement will be negatively associated with both parents' parenting quality, and with children's emotional adjustment. Retrospective reports indicate that an unequal childcare distribution is associated with suboptimal parenting behaviors: Increased authoritarian parenting in mothers, and disengaged parenting in fathers [@sabattiniRelationMothersFathers2004]. One reason might be that parents are less able to monitor each other when they are not equally involved [@mclanahanConsequencesFatherAbsence2014]. Furthermore, an unfair distribution of the demanding task childcare strains the interparental relationship, interferes with sensitive parenting, and undermines child well-being [@davies2006interparental; @deutschPaternalParticipationChild2001]. When care is distributed equally, mothers are more satisfied [@pharesFathersMothersInvolvement2009], and the family system is more harmonious. Consequently, balanced co-parenting is associated with greater parental warmth [@barnettInterdependenceParentingMothers2008].

@@

Bioecological theories posit that children require a certain amount of care; who provides it is somewhat flexible [see @bronfenbrennerEcologyHumanDevelopment1979]. We therefore hypothesize that greater joint care should positively predict children's emotional adjustment, $H_{6a}$. Further, it has been argued that children feel less supported when one parent is relatively uninvolved, than when both parents are equally invested in their well-being [@deutschPaternalParticipationChild2001]. We therefore predict that greater relative involvement will negatively predict emotional adjustment, $H_{6b}$.

@@

Interactions between involvement quality and quantity have largely been overlooked in the parenting literature [@jaffeeLifeFatherBenefits2003]. One notable exception is Feinberg's coparenting model [@feinbergInternalStructureEcological2003]. Coparenting refers to the ways parents relate to each other in the parenting role. Although the coparenting literature is typically invoked to describe child-rearing in divorced families, it also holds insights for intact families. According to Feinberg, the "division of labor" in childcare, analogous to our operationalization of relative involvement, affects qualitative child-rearing behaviors and children's adjustment directly, and moderates the link between these two constructs. Based on this insight, we might expect parents' quantitative involvement to moderate effects between qualitative parenting behaviors and emotional adjustment. As an example, parental anger might affect children's emotional adjustment differently when coming from a highly involved parent than from an absent parent. Conversely, children's poor emotional adjustment might elicit warm responses from hands-on parents, who are attuned to the reasons for these emotional troubles, than from out-of-touch parents, who might perceive maladjustment as a problem behavior that should be disciplined. Preliminary evidence indicates that interactions between involvement quality and quantity are indeed relevant. For instance, fathers' playful and caring involvement positively predicted preschoolers' emotional adjustment in school, but only when accompanied by a supportive coparenting relationship [@jiaTransactionalRelationsFather2012]. Furthermore, the quantity of paternal school-related involvement interacted with paternal warmth to positively predict adolescents' academic achievement and behavioral adjustment [@wangParentalInvolvementAfrican2014]. However, this literature is too limited to formulate specific hypotheses about moderating effects in the present study. We thus conduct an exploratory assessment of moderating effects of joint and relative quantitative involvement on the interplay between qualitative parenting behaviors and emotional adjustment.

@@

The present study examined associations between qualitative parenting behaviors and children's emotional adjustment, and whether these associations are moderated by quantitative involvement. We aimed to make several contributions. First, by including both mothers and fathers, we can tease apart their common and unique roles in children's emotional adjustment development. Second, by incorporating both qualitative and quantitative measures of involvement, we can examine interactions between qualitative and quantitative involvement, which have been largely overlooked in prior work. Third, by disentangling stable between-family differences from dynamic within-family predictive processes, we can establish at which level associations occur, and assess the within-family direction of effects between qualitative involvement and emotional adjustment. Finally, we examined child sex as a predictor of between-family differences and moderator of within-family effects.

@@

We employed data from the `r masked("K-cohort")` of the `r masked("Longitudinal Study of Australian Children (LSAC)")`, waves `r masked("2-5")`, which were collected by means of biannual home visits, and self-report mail questionnaires, from 2006-2012. Each wave was reviewed and approved by the
`r masked("Australian Institute of Family Studies Ethics Committee")`, which is a Human Research
Ethics Committee registered with the `r masked("National Health and Medical Research Council")` [@australianinstituteoffamilystudiesGrowingAustralia2020]. `r masked("LSAC")` employed a clustered sampling design to obtain a representative sample of the `r masked("Australian")` population. Complete details regarding the study design, sample, and procedure are available online [@australianinstituteoffamilystudiesGrowingAustralia2020].

@@

Scale descriptive statistics, attrition (which ranged from 7-14% across respondents), reliability, and intraclass correlations (ICCs) are reported in Supplementary Table 1; scale correlations are provided in Supplementary Table 2. The difference in fit between structurally invariant models and metrically invariant models was smaller than .01 on all conventional fit indices: RMSEA, CFI, and TLI (see Supplementary Table 3). We thus deemed it defensible to assume metric measurement invariance (i.e., to assume that the factor loadings of each construct are the same across time points and respondents), and calculated mean scores for all study variables.

@@

Responsiveness and inductive reasoning were measured using the six-item warmth (e.g., "How often do you express affection by hugging, kissing and holding this child?") and five-item inductive reasoning (e.g., "Talk it over and reason with this child when he/she misbehaved?") subscales of the Child Rearing Questionnaire [CRQ, @patersonAssociationBehaviouralAdjustment2001]. Consistency and negative parenting were measured using the five-item consistency (e.g., "How often is this child able to get out of punishment when he/she really sets his/her mind to it?", reverse coded) and six-item anger (e.g., "How often are you angry when you punish this child?") scales developed by the Canadian National Longitudinal Survey of Children and Youth [NLSCY; @statisticscanadaNationalLongitudinalSurvey2000]. These self-report instruments used five-point scales, ranging from "never or almost never" to "always or almost always".

@@

Measures of quantitative involvement varied from wave to wave, due to children's changing care needs [@wangATUSFathersInvolvement2009]. We used the first wave reports as an indicator of between-family differences in quantitative involvement. Both parents rated their involvement in four different child care activities on five item, five-point scales ranging from "once a day or more" to "not at all" [e.g., "In the past month how often did you help this child brush his/her teeth?"; @baxterFatheringAustraliaCouple2011].

@@

Emotional adjustment was measured using the reverse-coded five-item emotional distress subscale of the teacher-reported Strengths and Difficulties Questionnaire, on a three-point scale from "not true" to "certainly true" [e.g., "Often unhappy, downhearted or tearful"; @goodmanPsychometricPropertiesStrengths2001].
